---
title: "machin learning project"
author: "Leila"
date: "11/13/2021"
output: html_document
---

## Prerequisites

```{r}

library(lattice)
library(ggplot2)
library(caret)
library(rpart)
library(rpart.plot)
library(rattle)
library(randomForest)
library(RColorBrewer)

set.seed(12345)
```

## Data loading

```{r}
train <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", na.strings = c("", "NA"))
test <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", na.strings = c("", "NA"))
```

## Cleaning data
the columns that contain the near-zero variance data and a lot of NA values should be deleted because they don't have any significant effects on the model.

```{r}
## deleting near-zero variance data
near_zero <- nearZeroVar(train)
train <- train[,-near_zero] 
dim(train)

## deleting NA
na <- sapply(train, function(x) mean(is.na(x)))> 0.90
train <- train[, na==FALSE]
dim(train)

## deleting first fifth columns related to describing each person
train <- train[, -(1: 5)]
dim(train)
```

## Data partioning

```{r}
in_train <-createDataPartition(train$classe, p=.75, list=FALSE)
training <- train[in_train,]
testing <-train[-in_train,]
dim(training)
dim(testing)
```

## Random forest

```{r}
set.seed(123456)
model_rf <- train(classe ~ ., data = training, method = "rf")
pred_rf <- predict(model_rf, testing)
confusionMatrix(pred_rf, factor(testing$classe))
```

## Liner discriminant analysis

```{r}
set.seed(123456)
model_lda <- train(classe ~ ., data = training, method = "lda")
pred_lda <- predict(model_lda, testing)
confusionMatrix(factor(testing$classe),pred_lda)
```

## Recursive partitioning and regression trees

```{r}
set.seed(123456)
model_rpart <- train(classe ~ ., data = training, method = "rpart")
pred_rpart<- predict(model_rpart, testing)
confusionMatrix(factor(testing$classe), pred_rpart)
fancyRpartPlot(model_rpart$finalModel)
```

## Selecting the best model and applying on test data
Comparing the accuracy of the different tested models showed that high accuracy is related to Random Forest.

```{r}
predict_test <- as.data.frame(predict(model_rf, newdata = test))
predict_test
```




















